\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-20 forwards-and-backwards.tex
.

\begin{definition}[Model parameterization and assumptions]
    \label{def:model_setup}
    For each $\theta \ge 0$,
    let $\left(\eta^\theta(t)\right)_{t \ge 0}$
    be the continuous-time measure-valued Markov process described above
    with birth rates, establishment probabilities, and death rates
    determined as follows.
    Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
    are bounded, uniformly continuous functions
    from $\IR^d \times \IR_{\ge 0}$ to $\IR_{\ge 0}$.
    Furthermore, suppose $r$ takes values in $[0, 1]$
    and has uniformly bounded second derivatives,
    and that $(x, m) \mapsto m^2 \gamma(x, m)$ is unformly bounded.
    Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
    with $\epsilon_r < \epsilon_\gamma - 1/\theta$.
    Suppose $F: \IR^d \times \IR_{\ge 0} \to \IR$ is bounded and uniformly continous
    and define
    \begin{align}
        \mu(x, \eta)
        =
        r(x, \smooth{r} \eta(x))
        \gamma(\smooth{\gamma} \eta(x))
        - \frac{1}{\theta}
        F(x, \smooth{F} \eta(x)) .
    \end{align}
    Let $N$ be a positive number determined by $\theta$.
    When the state of the process is $\eta$,
    offspring of mass $1/N$ are produced at location $x$
    with rate $\theta \gamma(x, \smooth{\gamma} \eta^{\theta}(x)) N \eta^{\theta}_s(dx)$
    and adults die at location $x$ with rate $\theta \mu(x, \eta) N \eta^{\theta}_s(dx)$,
    (i.e., at rates $\theta \gamma$ and $\theta \mu$ per capita respectively).
    Offspring disperse to a location offset from the parent's location
    by an independent Gaussian with mean $\meanq/\theta$
    and covariance matrix $\covq / \theta$;
    an offspring at $x$ establishes instantaneously
    with probability $r(x, \smooth{r} \eta^{\theta}(x))$, or else dies.
\end{definition}

@@

Putting these together,
the level $u$ of an individual at $x$
evolves according to
$$
\dot{u}
=
    - \frac{u}{\lambda} \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \mu(x,\eta) u .
$$
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
(This differs from $F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$ defined above
because $b$ contains the adjustment to fecundity obtained by
averaging over establishment probabilities in the locations that offspring disperse to.)
\comment{TODO: make this explanation better.}
We can then rewrite the differential equation 
governing the how each individual's level evolves as
\begin{align}
\dot{u}
    &=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    \left\{
        -\frac{u}{\lambda}\left(\lambda - u\right)
        + u
    \right\}
    -
    b(x,\eta)u
    \nonumber \\
    &=
    \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
    -
    b(x, \eta) u / \theta
    . \label{differential equation for level}
\end{align}

@@

First consider the terms in the generator that come from birth events.
Recall that when a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align}
f(\lp) \nonumber
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{\lambda} \gamma(x, \eta)
    \int_u^\lambda
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    \\
    \begin{split} \label{eqn:birth_generator}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, v_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.

@@

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
    \label{Def: Rescaled Lookdown Models}
We define $(\lp^{\theta}_t)_{t \geq 0}$
to be the (unique) solution
    \comment{(I'm not sure why unique is in parentheses here?)}
to the Martingale Problem
with initial condition $\lp_0 \in \mathcal{M}_F(\mathbb{R}^d \times [0,N])$
and generator $A^{\theta}$ that satisfies
\begin{equation*}
\begin{split}
& A^{\theta}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac 12 N^{-1}\int_u^N g(x,u_1) du_1
            \times
            \frac{
                \theta \int_{\mathbb{R}^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        N^{-1}\theta\int_u^N
        \int_{\mathbb{R}^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        \theta N^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) u^2 -b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation*}
for all smooth functions $g \in C^{2}_{0}(\mathbb{R}^d)$,
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$ is the spatial pushforward of $\lp$,
as before.
\end{definition}

@@

\begin{definition}
    \label{def: MP definition of limit}
Let $F \in \mathcal{C}^{2}(\mathbb{R})$ and $f \in \mathcal{C}_{0}(\mathbb{R}^d)$.
We define $\Pgen$ to be a generator
acting on test functions of the form
$F_f(\eta):=F (\langle f, \eta \rangle)$
such that 
\begin{equation}
\begin{aligned}
\Pgen^{\infty} F_f(\eta):=& F'(\langle f, \eta \rangle)
                   \times \left\{
                   \big\langle
                        \gamma(x, \smooth{\gamma} \eta(x))
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \smooth{r} \eta(\cdot))
                            \right)(x)
                    +
                    f(x)
                        F(x, \smooth{F} \eta(x)),
                        \eta(dx)
                    \big\rangle
                   \right\}\\
                &+ \frac{1}{2}F''(\langle f, \eta \rangle)
                  \times \left\{
                  \alpha \big\langle
                    \gamma\left( x, \smooth{\gamma} \eta(x) \right)
                    r\left(x, \smooth{r} \eta(x) \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle 
                  \right\}.
\end{aligned}    
\end{equation}
We define $(\eta^{\infty}(t))_{t \geq 0}$ to be the solution to the Martingale Problem $(\Pgen^{\infty}, \eta_0)$, i.e. 
for all $F \in \mathcal{C}^{2}(\mathbb{R})$
and $f \in \mathcal{C}_{0}(\mathbb{R}^d)$,
$$M_t:=F_f(\eta^{\infty}_t)-F_f(\eta_0)
-\int_{0}^{t}\Pgen^{\infty}F_f(\eta^{\infty}_s)ds$$
is a martingale.
\end{definition}
\begin{remark}
Taking $F(x)=x$ and $F(x)=x^2$, we can recover Equation \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}.
\end{remark}
\begin{lemma}[Convergence of generators for the population process]
As $\theta \to \infty$,
$(\eta^{\theta}_t)_{t \geq 0}$ converges in $\mathcal{D}_{[0,\infty)}(\mathcal{M}_f(\mathbb{R}^d))$
to $(\eta^{\infty}(t))_{t \geq 0}$.
\end{lemma}
\begin{proof}
We use Theorem 4.8.10 in \cite{EK}.
First of all, the set of functions
$\{F_f(\eta):= F(\langle f, \eta \rangle ),~
F \in \mathcal{C}^{2}(\mathbb{R}), ~
f \in \mathcal{C}_{0}(\mathbb{R}^d)\}$
is a dense and convergence-determining 
subset of functions in $\mathcal{M}_f(\mathbb{R}^d)$.
Therefore, it suffices to show that for any $t>0$,
\begin{equation}
    \label{eq: Convergence Condition}
\lim_{\theta \to \infty}
\mathbb{E}\left[
\left(
F_f(\eta^{\theta}_{t+s})-F_f(\eta^{\theta}_t)
-\int_{t}^{t+s}\Pgen^{\infty}F_f(\eta^{\theta}_s)ds
\right)
\prod_{i=1}^{k}\langle h_i,\eta^{\theta}_{t_i} \rangle
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+s$,
and $h_1,...,h_k \in \mathcal{C}_{0}(\mathbb{R}^d)$.

@@

>>>>

\begin{definition}[Model parameterization and assumptions]
    \label{def:model_setup}
    For each $\theta \ge 0$,
    let $\left(\eta^\theta(t)\right)_{t \ge 0}$
    be the continuous-time measure-valued Markov process described above
    with birth rates, establishment probabilities, and death rates
    determined as follows.
    Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
    are bounded, uniformly continuous functions
    from $\IR^d \times \IR_{\ge 0}$ to $\IR_{\ge 0}$.
    Furthermore, suppose $r$ takes values in $[0, 1]$
    and has uniformly bounded second derivatives,
    and that $(x, m) \mapsto m^2 \gamma(x, m)$ is unformly bounded.
    Fix nonnegative values $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
    with $\epsilon_r < \epsilon_\gamma - 1/\theta$.
    Suppose $F: \IR^d \times \IR_{\ge 0} \to \IR$ is bounded and uniformly continous
    and define
    \begin{align} \label{eqn:mu_defn}
        \mu(x, \eta)
        =
        r(x, \smooth{r} \eta(x))
        \gamma(\smooth{\gamma} \eta(x))
        - \frac{1}{\theta}
        F(x, \smooth{F} \eta(x)) .
    \end{align}
    Let $N$ be a positive number determined by $\theta$.
    When the state of the process is $\eta$,
    offspring of mass $1/N$ are produced at location $x$
    with rate $\theta \gamma(x, \smooth{\gamma} \eta^{\theta}(x)) N \eta^{\theta}_s(dx)$
    and adults die at location $x$ with rate $\theta \mu(x, \eta) N \eta^{\theta}_s(dx)$,
    (i.e., at rates $\theta \gamma$ and $\theta \mu$ per capita respectively).
    Offspring disperse to a location offset from the parent's location
    by an independent Gaussian with mean $\meanq/\theta$
    and covariance matrix $\covq / \theta$;
    an offspring at $x$ establishes instantaneously
    with probability $r(x, \smooth{r} \eta^{\theta}(x))$, or else dies.
\end{definition}

@@

Putting these together,
the level $u$ of an individual at $x$ in $\lp$ evolves according to:
$$
    \dot u
    =
    - \frac{\theta}{\lambda} u \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \theta \mu(x,\eta) u .
$$
The factor of $\theta$ appears here because we have rescaled time.
We shall write 
$$
    b(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta)
    \right) ,
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
Recall from equation \eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu(x,\eta))$,
and so we will have that
\begin{align*}
b(x, \eta)
&=
    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) q(x, dy)
    + F(x, \eta) \\
&\to
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) \qquad \text{as } \theta \to \infty .
\end{align*}

We can then rewrite the differential equation 
governing how the level of each individual
in $\lp$ evolves as
\begin{align}
\dot{u}
    &=
    \theta \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    \left\{
        -\frac{u}{\lambda}\left(\lambda - u\right)
        + u
    \right\}
    -
    b(x,\eta) u
    \nonumber \\
    &=
    \frac{\theta}{\lambda} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
    -
    b(x, \eta) u
    . \label{differential equation for level}
\end{align}
(The factor of $\theta$ appears because of the rescaling of time for $\lp$.)

@@

First consider the terms in the generator that come from birth events.
Recall that when a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align}
f(\lp) \nonumber
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{\lambda} \gamma(x, \eta)
    \int_u^\lambda
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    \\
    \begin{split} \label{eqn:birth_generator}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.
We've written it in this form because the two pieces
each naturally converge to separate terns in the limit.

@@

% % % % % % % %
\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
We define $(\lp^{\theta}_t)_{t \geq 0}$
to be the (unique) solution
    \comment{(I'm not sure why unique is in parentheses here?)}
to the Martingale Problem
with initial condition $\lp_0 \in \mathcal{M}_F(\mathbb{R}^d \times [0,N])$
and generator $A^{\theta}$ that satisfies
\begin{equation*}
\begin{split}
& A^{\theta}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac 12 N^{-1}\int_u^N g(x,u_1) du_1
            \times
            \frac{
                \theta \int_{\mathbb{R}^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        N^{-1}\theta\int_u^N
        \int_{\mathbb{R}^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        \theta N^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) u^2 -b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation*}
for all smooth functions $g \in C^{2}_{0}(\mathbb{R}^d)$,
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$ is the spatial pushforward of $\lp$,
as before.
\end{definition}

@@

\begin{definition}
    \label{def: MP definition of limit}
Let $F \in \mathcal{C}^{2}(\mathbb{R})$ and $f \in \mathcal{C}_{0}(\mathbb{R}^d)$.
We define $\Pgen$ to be a generator
acting on test functions of the form
$F_f(\eta):=F (\langle f, \eta \rangle)$
such that 
\begin{equation}
\begin{aligned}
\Pgen^{\infty} F_f(\eta):=& F'(\langle f, \eta \rangle)
                   \times \left\{
                   \big\langle
                        \gamma(x, \smooth{\gamma} \eta(x))
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \smooth{r} \eta(\cdot))
                            \right)(x)
                    +
                    f(x)
                        F(x, \smooth{F} \eta(x)),
                        \eta(dx)
                    \big\rangle
                   \right\}\\
                &+ \frac{1}{2}F''(\langle f, \eta \rangle)
                  \times \left\{
                  \alpha \big\langle
                    \gamma\left( x, \smooth{\gamma} \eta(x) \right)
                    r\left(x, \smooth{r} \eta(x) \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle 
                  \right\}.
\end{aligned}    
\end{equation}
We define $(\eta^{\infty}(t))_{t \geq 0}$ to be the solution to the Martingale Problem $(\Pgen^{\infty}, \eta_0)$, i.e. 
for all $F \in \mathcal{C}^{2}(\mathbb{R})$
and $f \in \mathcal{C}_{0}(\mathbb{R}^d)$,
$$M_t:=F_f(\eta^{\infty}_t)-F_f(\eta_0)
-\int_{0}^{t}\Pgen^{\infty}F_f(\eta^{\infty}_s)ds$$
is a martingale.
\end{definition}
\begin{remark}
Taking $F(x)=x$ and $F(x)=x^2$, we can recover Equation \eqref{eqn:mgale_problem} and \eqref{eqn:mgale_variation}.
\end{remark}

\begin{lemma}[Convergence of generators for the population process]
As $\theta \to \infty$,
$(\eta^{\theta}_t)_{t \geq 0}$ converges in $\mathcal{D}_{[0,\infty)}(\mathcal{M}_f(\mathbb{R}^d))$
to $(\eta^{\infty}(t))_{t \geq 0}$.
\end{lemma}
\begin{proof}
We use Theorem 4.8.10 in \cite{EK}.
First of all, the set of functions
$\{F_f(\eta):= F(\langle f, \eta \rangle ),~
F \in \mathcal{C}^{2}(\mathbb{R}), ~
f \in \mathcal{C}_{0}(\mathbb{R}^d)\}$
is a dense and convergence-determining 
subset of functions in $\mathcal{M}_f(\mathbb{R}^d)$.
Therefore, it suffices to show that for any $t>0$,
\begin{equation}
    \label{eq: Convergence Condition}
\lim_{\theta \to \infty}
\mathbb{E}\left[
\left(
F_f(\eta^{\theta}_{t+s})-F_f(\eta^{\theta}_t)
-\int_{t}^{t+s}\Pgen^{\infty}F_f(\eta^{\theta}_s)ds
\right)
\prod_{i=1}^{k}\langle h_i,\eta^{\theta}_{t_i} \rangle
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+s$,
and $h_1,...,h_k \in \mathcal{C}_{0}(\mathbb{R}^d)$.

@@

\comment{Here's an informal writeup of this convergence;
mostly I'm doing it to make sure I've got the right stuff above.}

In Definition~\ref{defn:lookdown_mgale} we saw that the generator
of $(\lp_t)_{t \ge 0}$ is the sum of two parts,
equations \eqref{eqn:birth_generator} and \eqref{eqn:level_generator}.
Assume that there is a $u_*$ such that $g(x, u) = 1$ for all $u \ge u_*$.
We consider these two in turn.

First, the contribution from births is
\begin{align*}
\begin{split}
f(\lp)
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align*}
We have that
\begin{align*}
    \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
    \to
    \DG(gr)(x) \qquad \text{as }\theta \to \infty,
\end{align*}
i.e., to $\DG\left(g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)$.
Furthermore, by our assumption on $g$,
$\int_u^\lambda g(x, u_1) du_1 / \lambda \to 1$ for all $x$ and $u$,
so the first term here converges to
\begin{align*}
    2 \gamma(x, \eta)
        \frac{1}{2}
        \frac{
            \DG\left( g(\cdot, u) r(\cdot, \smooth{r}\eta(\cdot))\right)(x)
        }{
            g(x, u)
        } .
\end{align*}
This corresponds to each particle moving according to the motion induced by
$g \mapsto \gamma \DG(gr)$.
\comment{TODO: explain the $r$-tilted $\DG$ motion somewhere.}
As for the second term,
since $\int g(y, u_1) q(x, dy) \to g(x, u_1)$
and $\theta/\lambda \to \delta$,
it converges to
\begin{align*}
    \delta
    \int_u^\infty
    \left( g(x, u_1) - 1 \right)
    r(x, \eta) .
\end{align*}
This corresponds to births at higher levels at rate $\gamma r$.
